---
title: "OR538 Final Project - ARIMA Modeling"
output: html_notebook
---

# Modeling Paycheck Inc.

In this section we will be applying AR, MA, ARMA and ARIMA modeling for the PayCheck Inc. security and evaluating the models by using Ljung-Box Test on the residuals of each model.

## AR Modeling
Key points:
* Stationarity implies that the amount of variance is equal to the amount of feedback ($|\phi|<1$)
* $Y_t=(1-\phi)\mu+\phi{Y_{t-1}}+\epsilon_t$
* $Y_t=\beta_0+\beta_1{Y_t}+\epsilon_t$
* For model evalutation using Ljung-Box needs the fitdf parameter to be adjusted and its value depends on the AR model. For instance, for an AR(1) model the $fitdf=1$
* In the test a p-value of 0 indicates that at least one of the first **df** lags present an autocorrelation that is not zero

For each results plot residuals against time, ACF for the residuals and q-q plot using normal distribution for the errors (i.e. qqnorn).

any significant autocorrelation in the residuals implies that the model did not fit well.

One typical problem of AR(p) models is that they generally need a large number of p to fit the data set.

============


```{r}
pacman::p_load("forecast") # Library needed for auto.arima

AR0 <- arima(df_log_ret[,"PAYX"], order = c(0,0,0))
AR1 <- arima(df_log_ret[,"PAYX"], order = c(1,0,0))
AR2 <- arima(df_log_ret[,"PAYX"], order = c(2,0,0))
AR3 <- arima(df_log_ret[,"PAYX"], order = c(3,0,0))

AR0Res <- c("0", AR0$aic)
AR1Res <- c("1", AR1$aic)
AR2Res <- c("2", AR2$aic)
AR3Res <- c("3", AR3$aic)

df_AR_results <- as.data.frame(rbind(AR0Res, AR1Res, AR2Res, AR3Res))
colnames(df_AR_results) <- c("p", "AIC")

df_AR_results
```
**Table 2:** AR Model Results
 
Table **2** allow us to observe that the best result is obtained when we use an AR(0) model.

# MA Modeling

```{r}
MA0 <- arima(df_log_ret[,"PAYX"], order = c(0,0,0))
MA1 <- arima(df_log_ret[,"PAYX"], order = c(0,0,1))
MA2 <- arima(df_log_ret[,"PAYX"], order = c(0,0,2))
MA3 <- arima(df_log_ret[,"PAYX"], order = c(0,0,3))

MA0Res <- c("0", MA0$aic)
MA1Res <- c("1", MA1$aic)
MA2Res <- c("2", MA2$aic)
MA3Res <- c("3", MA3$aic)

df_MA_results <- as.data.frame(rbind(MA0Res, MA1Res, MA2Res, MA3Res))
colnames(df_MA_results) <- c("q", "AIC")

df_MA_results
```
**Table 3:** MA Model Results
 
Table **3** allow us to observe that the best result is obtained when we use an MA(0) model.


# ARMA Modeling

```{r}
ARMA00 <- arima(df_log_ret[,"PAYX"], order = c(0,0,0))
ARMA01 <- arima(df_log_ret[,"PAYX"], order = c(0,0,1))
ARMA02 <- arima(df_log_ret[,"PAYX"], order = c(0,0,2))
ARMA03 <- arima(df_log_ret[,"PAYX"], order = c(0,0,3))

ARMA10 <- arima(df_log_ret[,"PAYX"], order = c(1,0,0))
ARMA11 <- arima(df_log_ret[,"PAYX"], order = c(1,0,1))
ARMA12 <- arima(df_log_ret[,"PAYX"], order = c(1,0,2))
ARMA13 <- arima(df_log_ret[,"PAYX"], order = c(1,0,3))

ARMA20 <- arima(df_log_ret[,"PAYX"], order = c(2,0,0))
ARMA21 <- arima(df_log_ret[,"PAYX"], order = c(2,0,1))
ARMA22 <- arima(df_log_ret[,"PAYX"], order = c(2,0,2))
ARMA23 <- arima(df_log_ret[,"PAYX"], order = c(2,0,3))

ARMA30 <- arima(df_log_ret[,"PAYX"], order = c(3,0,0))
ARMA31 <- arima(df_log_ret[,"PAYX"], order = c(3,0,1))
ARMA32 <- arima(df_log_ret[,"PAYX"], order = c(3,0,2))
ARMA33 <- arima(df_log_ret[,"PAYX"], order = c(3,0,3))

ARMA00Res <- c("0", "0", ARMA00$aic)
ARMA01Res <- c("0", "1", ARMA01$aic)
ARMA02Res <- c("0", "2", ARMA02$aic)
ARMA03Res <- c("0", "3", ARMA03$aic)

ARMA10Res <- c("1", "0", ARMA10$aic)
ARMA11Res <- c("1", "1", ARMA11$aic)
ARMA12Res <- c("1", "2", ARMA12$aic)
ARMA13Res <- c("1", "3", ARMA13$aic)

ARMA20Res <- c("2", "0", ARMA20$aic)
ARMA21Res <- c("2", "1", ARMA21$aic)
ARMA22Res <- c("2", "2", ARMA22$aic)
ARMA23Res <- c("2", "3", ARMA23$aic)

ARMA30Res <- c("3", "0", ARMA30$aic)
ARMA31Res <- c("3", "1", ARMA31$aic)
ARMA32Res <- c("3", "2", ARMA32$aic)
ARMA33Res <- c("3", "3", ARMA33$aic)

df_ARMA_results <- as.data.frame(rbind(ARMA00Res, ARMA01Res, ARMA02Res, ARMA03Res,
                                       ARMA10Res, ARMA11Res, ARMA12Res, ARMA13Res,
                                       ARMA20Res, ARMA21Res, ARMA22Res, ARMA23Res,
                                       ARMA30Res, ARMA31Res, ARMA32Res, ARMA33Res))
colnames(df_ARMA_results) <- c("p", "q", "AIC")

df_ARMA_results
```
**Table 4:** ARMA Model Results
 
Table **4** allow us to observe that the best result is obtained when we use an ARMA(0,0) model.

```{r}
ARIMAautoAic <- auto.arima(df_log_ret[,"PAYX"], max.p = 20, max.q = 20, max.d = 20, ic = "aic")

ARIMAautoAic
```

# Conclusion
Manually performing ARIMA models for each stock could be a tedious task and require a significant amount of effort to perform. Since the end goal of the project is to get as close as possible to generate an optimal portfolio for investment we will focus the next sections on using `auto.arima` function with each stock and evaluate the residuals for each stock.
