---
title: "Final Project"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

The first step is to create the dataset by pulling values and generating
the rate of change of interest (i.e., log-returns)

In this case we are going to use the stock prices for two companies in the
information technology industry in particular the Data & Processing and
outsourced services (i.e., Paycheck inc. and PayPal), one company in the
Consumer Staples industry (i.e., Walmart), one market (i.e., S&P 500) and
the risk-free rates for 2019 provided by the US Department of Treasury.



```{r}
pacman::p_load(quantmod, tseries, qmao)
# Load prices from Yahoo finance
# PAYX = PayCheck Inc.
# PYPL = PayPal
# WMT = WALMART
# SP = S&P 500

symbols <- c("PAYX", "PYPL", "WMT", "SP")

getSymbols(symbols, from = "2017-11-23", to = "2019-11-23")


df_adj <- data.frame(PAYX=PAYX$PAYX.Adjusted,
                     PYPL=PYPL$PYPL.Adjusted,
                     WMT=WMT$WMT.Adjusted,
                     SP=SP$SP.Adjusted)
colnames(df_adj) <- symbols

df_ret <- data.frame(PAYX=diff(PAYX$PAYX.Adjusted),
                     PYPL=diff(PYPL$PYPL.Adjusted),
                     WMT=diff(WMT$WMT.Adjusted),
                     SP=diff(SP$SP.Adjusted))
colnames(df_ret) <- symbols

df_log_ret <- data.frame(PAYX=diff(log(PAYX$PAYX.Adjusted)),
                         PYPL=diff(log(PYPL$PYPL.Adjusted)),
                         WMT=diff(log(WMT$WMT.Adjusted)),
                         SP=diff(log(SP$SP.Adjusted)))
colnames(df_log_ret) <- symbols

#df_log_ret <- diff(log(ts(df_adj)))

#Res_Raw <- getSymbols('CPIAUCNS', from = "2011-11-23", to = "2019-11-23", src='FRED')



# AAPL_Adjusted <- AAPL$AAPL.Adjusted
#auto.arima(TbGdpPi[,1],max.P=0,max.Q=0,ic="aic")

# ccf(df_log_ret, lwd=3, ylab= "CCF", main="test")

```

First question was to determine which of the 2 most common attributes to use 
(i.e., returns or log returns))

```{r}
pdf("Figure1-price.pdf",height=7,width=8)
plot(as.xts(df_adj), main="Adjusted Price", legend.loc = "topleft")
graphics.off()

plot(as.xts(df_adj), main="Adjusted Price", legend.loc = "topleft")
```
**Figure 1.** Time-Series plot for the Adjusted Price of PayCheck Inc, PayPal,
Walmart and SP.

At first glance we can observe quite some difference between the price values
for the individual stocks and the market price.

```{r}
pdf("Figure2-returns.pdf",height=7,width=8)
plot(as.xts(df_ret), main="Returns", legend.loc = "topleft")
graphics.off()

plot(as.xts(df_ret), main="Returns", legend.loc = "topleft")
```
**Figure 2.** Time-Series plot for the returns for PayCheck Inc, PayPal,
Walmart and SP.

The time-series starts to show a mean-reversion condition which is an indicator
used to identify distributional invariance over time, or stationarity. However,
it seems that we can't observe well the S&P 500 returns (i.e., blue lines).

**Key Concept:** An important consideration when modeling time series is the
principle of parsimony. This principle refers to representing the systematic
structure of the series with as few parameters as possible. Essentially, this
means simpler representations of a time series process are more desirable than
more complex ones if both are adequate. (Chapter 2. Autoregressive Integrated
Moving Average Models) (http://www.scausa.com/tsfbook.php#:~:targetText=An%20important%20consideration%20when%20modeling,as%20few%20parameters%20as%20possible.&targetText=This%20chapter%20examines%20seasonal%20time,are%20useful%20for%20such%20data)



```{r}
pdf("Figure3-log-returns.pdf",height=7,width=8)
plot(as.xts(df_log_ret), main="Log Returns", legend.loc = "topleft")
graphics.off()

plot(as.xts(df_log_ret), main="Log Returns", legend.loc = "topleft")
```
**Figure 3.** Time-Series plot for the log returns for PayCheck Inc, Paypal,
Walmart and SP.

The time-series continues to show the mean-reversion condition and on top the
color code for each security seems to be more evently distributed, including
S&P 500.

Since the log-returns seems to show better the parsimony principle; we will
continue the analysis using the log-return transformation.


# Are the time series stationary?

```{r}
pdf("Figure4-ACF-lag-default.pdf",height=7,width=8)
par(mfrow=c(2,2))
acf(df_log_ret[,"PAYX"], main="PayCheck Inc.", na.action = na.pass)
acf(df_log_ret[,"PYPL"], main="PayPal", na.action = na.pass)
acf(df_log_ret[,"WMT"], main="Walmart", na.action = na.pass)
acf(df_log_ret[,"SP"], main="S&P 500", na.action = na.pass)
graphics.off()

par(mfrow=c(2,2))
acf(df_log_ret[,"PAYX"], main="PayCheck Inc.", na.action = na.pass)
acf(df_log_ret[,"PYPL"], main="PayPal", na.action = na.pass)
acf(df_log_ret[,"WMT"], main="Walmart", na.action = na.pass)
acf(df_log_ret[,"SP"], main="S&P 500", na.action = na.pass)
```

The auto corrleation function applied to each security decays to zero, clearly 
indicating that the log-return series are stationary. However, PayPal security
and S&P 500 present ocasions where the autocorrelation is outside the bounds
which is an indication of non-stationary conditions.

In particular the autocorrelation that crossed the test bound for S&P 500
ocurred with a lag close to 15 and the autocorrelation that crossed the test
bound for PayPal was at a lag of 20. This allow us to generalize by indicating
that non of the securitiues autocorrelations at low lags (i.e., lower than 12)
are outside the test bounds.

**Reminder:**
 * If the autocorrelation cross the test bound the hypothesis is rejected
 * If the autocorrelation does not cross the test bound the hypothesis is not
 rejected.

```{r}
pdf("Figure5-ACF-Lag-12.pdf",height=7,width=8)
par(mfrow=c(2,2))
acf(df_log_ret[,"PAYX"], lag.max = 12, main="PayCheck Inc.; Max Lag = 12", na.action = na.pass)
acf(df_log_ret[,"PYPL"], lag.max = 12, main="PayPal; Max Lag = 12", na.action = na.pass)
acf(df_log_ret[,"WMT"], lag.max = 12, main="Walmart; Max Lag = 12", na.action = na.pass)
acf(df_log_ret[,"SP"], lag.max = 12, main="S&P 500; Max Lag = 12", na.action = na.pass)
graphics.off()

par(mfrow=c(2,2))
acf(df_log_ret[,"PAYX"], lag.max = 12, main="PayCheck Inc.; Max Lag = 12", na.action = na.pass)
acf(df_log_ret[,"PYPL"], lag.max = 12, main="PayPal; Max Lag = 12", na.action = na.pass)
acf(df_log_ret[,"WMT"], lag.max = 12, main="Walmart; Max Lag = 12", na.action = na.pass)
acf(df_log_ret[,"SP"], lag.max = 12, main="S&P 500; Max Lag = 12", na.action = na.pass)
```

The ACF plot for the securities were obtained with a max lag value of 12 and
we can observe that in this case; we have no autocorrelation crossing the 
test bound.

# Fitting the Models with AR(1)

**Note:** all the models are using the Maximum likelihood as the primary estimation
method.

```{r}
fitPayxAr1 <- arima(df_log_ret[,"PAYX"], order = c(1,0,0))
fitPyplAr1 <- arima(df_log_ret[,"PYPL"], order = c(1,0,0))
fitWmtAr1 <- arima(df_log_ret[,"WMT"], order = c(1,0,0))
fitSpAr1 <- arima(df_log_ret[,"SP"], order = c(1,0,0))

Box.test(residuals(fitPayxAr1), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPyplAr1), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitWmtAr1), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitSpAr1), lag = 5, type = "Ljung-Box", fitdf = 1)
```

# Forecasting PayCheck AR Model

```{r}
pred.payx <- predict(fitPayxAr1, n.ahead = 100, se.fit = TRUE)

t1 = (503-99):503
t2 = 504:(504+49+50)
day_range = seq(-100,100,1)

xts1 <- as.data.frame(df_log_ret)

plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx$pred - 2*pred.payx$se)
lines(day_range[101:200], pred.payx$pred + 2*pred.payx$se)
legend(50,-0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

pdf("Figure6-Forecast-Paycheck_AR1.pdf",height=7,width=8)
plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx$pred - 2*pred.payx$se)
lines(day_range[101:200], pred.payx$pred + 2*pred.payx$se)
legend(50,-0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
graphics.off()
```

 
# Forecasting Paypal AR(1) Model
 
```{r}
pred.pypl <- predict(fitPyplAr1, n.ahead = 100, se.fit = TRUE)

plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl$pred - 2*pred.pypl$se)
lines(day_range[101:200], pred.pypl$pred + 2*pred.pypl$se)
legend(50,0.085,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

pdf("Figure7-Forecast-Paypal_AR1.pdf",height=7,width=8)
plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl$pred - 2*pred.pypl$se)
lines(day_range[101:200], pred.pypl$pred + 2*pred.pypl$se)
legend(50,-0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
graphics.off()
```
 
 
# Forecasting Walmart AR(1) Model
 
```{r}
pred.wmt <- predict(fitWmtAr1, n.ahead = 100, se.fit = TRUE)

plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt$pred - 2*pred.wmt$se)
lines(day_range[101:200], pred.wmt$pred + 2*pred.wmt$se)
legend(50,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

pdf("Figure8-Forecast-Walmart_AR1.pdf",height=7,width=8)
plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt$pred - 2*pred.wmt$se)
lines(day_range[101:200], pred.wmt$pred + 2*pred.wmt$se)
legend(50,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
graphics.off()
```
 
 
# Forecasting S&P 500 AR(1) Model
 
```{r}
pred.sp <- predict(fitSpAr1, n.ahead = 100, se.fit = TRUE)

plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp$pred - 2*pred.sp$se)
lines(day_range[101:200], pred.sp$pred + 2*pred.sp$se)
legend(50,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

pdf("Figure9-Forecast-SP500_AR1.pdf",height=7,width=8)
plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp$pred - 2*pred.sp$se)
lines(day_range[101:200], pred.sp$pred + 2*pred.sp$se)
legend(50,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
graphics.off()
```


```{r}
pdf("Figure10-Forecast-All_AR1.pdf",height=7,width=8)
par(mfrow=c(2,2))
#
plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx$pred - 2*pred.payx$se)
lines(day_range[101:200], pred.payx$pred + 2*pred.payx$se)
legend(25,0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl$pred - 2*pred.pypl$se)
lines(day_range[101:200], pred.pypl$pred + 2*pred.pypl$se)
legend(25,0.09,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt$pred - 2*pred.wmt$se)
lines(day_range[101:200], pred.wmt$pred + 2*pred.wmt$se)
legend(25,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp$pred - 2*pred.sp$se)
lines(day_range[101:200], pred.sp$pred + 2*pred.sp$se)
legend(25,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
graphics.off()
```
 

# Fitting the Models AR(p)

```{r}
pacman::p_load("forecast")

fitPayxArpBic <- auto.arima(df_log_ret[,"PAYX"], max.p = 20, max.q = 0, ic = "bic")
fitPyplArpBic <- auto.arima(df_log_ret[,"PYPL"], max.p = 20, max.q = 0, ic = "bic")
fitWmtArpBic <- auto.arima(df_log_ret[,"WMT"], max.p = 20, max.q = 0, ic = "bic")
fitSpArpBic <- auto.arima(df_log_ret[,"SP"], max.p = 20, max.q = 0, ic = "bic")

fitPayxArpAic <- auto.arima(df_log_ret[,"PAYX"], max.p = 20, max.q = 0, ic = "bic")
fitPyplArpAic <- auto.arima(df_log_ret[,"PYPL"], max.p = 20, max.q = 0, ic = "bic")
fitWmtArpAic <- auto.arima(df_log_ret[,"WMT"], max.p = 20, max.q = 0, ic = "bic")
fitSpArpAic <- auto.arima(df_log_ret[,"SP"], max.p = 20, max.q = 0, ic = "bic")

Box.test(residuals(fitPayxArpBic), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPayxArpAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitPyplArpBic), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPyplArpAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitWmtArpBic), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitWmtArpAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitSpArpBic), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitSpArpAic), lag = 5, type = "Ljung-Box", fitdf = 1)
```

Below is the script to generate the forecast for all four securities only using AR(p) model and optimizing AIC.

```{r}
pred.payx.arp.aic <- predict(fitPayxArpAic, n.ahead = 100, se.fit = TRUE)
pred.pypl.arp.aic <- predict(fitPyplArpAic, n.ahead = 100, se.fit = TRUE)
pred.wmt.arp.aic <- predict(fitWmtArpAic, n.ahead = 100, se.fit = TRUE)
pred.sp.arp.aic <- predict(fitSpArpAic, n.ahead = 100, se.fit = TRUE)

pdf("Figure11-Forecast-All_ARp.pdf",height=7,width=8)
par(mfrow=c(2,2))
#
plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx.arp.aic$pred - 2*pred.payx.arp.aic$se)
lines(day_range[101:200], pred.payx.arp.aic$pred + 2*pred.payx.arp.aic$se)
legend(25,0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl.arp.aic$pred - 2*pred.pypl.arp.aic$se)
lines(day_range[101:200], pred.pypl.arp.aic$pred + 2*pred.pypl.arp.aic$se)
legend(25,0.09,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt.arp.aic$pred - 2*pred.wmt.arp.aic$se)
lines(day_range[101:200], pred.wmt.arp.aic$pred + 2*pred.wmt.arp.aic$se)
legend(25,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp.arp.aic$pred - 2*pred.sp.arp.aic$se)
lines(day_range[101:200], pred.sp.arp.aic$pred + 2*pred.sp.arp.aic$se)
legend(25,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
graphics.off()
```

Below is the script to generate a figure that would allow us to compare the forecast created
using the predictions for the AR(1) model and AR(p) model for the PayCheck Inc. and Paypal securities.

```{r}
pdf("Figure12-Forecast-Payx-Pypl_AR1_ARp.pdf",height=7,width=8)
par(mfrow=c(2,2))
#
plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast - AR(p)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx.arp.aic$pred - 2*pred.payx.arp.aic$se)
lines(day_range[101:200], pred.payx.arp.aic$pred + 2*pred.payx.arp.aic$se)
legend(25,0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast - AR(1)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx$pred - 2*pred.payx$se)
lines(day_range[101:200], pred.payx$pred + 2*pred.payx$se)
legend(25,0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast - AR(p)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl.arp.aic$pred - 2*pred.pypl.arp.aic$se)
lines(day_range[101:200], pred.pypl.arp.aic$pred + 2*pred.pypl.arp.aic$se)
legend(25,0.09,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast - AR(1)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl$pred - 2*pred.pypl$se)
lines(day_range[101:200], pred.pypl$pred + 2*pred.pypl$se)
legend(25,0.09,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
graphics.off()
```

Below is the script to generate a figure that would allow us to compare the forecast created
using the predictions for the AR(1) model and AR(p) model for the Walmart and S&P 500 securities.

```{r}
pdf("Figure13-Forecast-Wmt-Sp_AR1_ARp.pdf",height=7,width=8)
par(mfrow=c(2,2))
#
plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast - AR(p)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt.arp.aic$pred - 2*pred.wmt.arp.aic$se)
lines(day_range[101:200], pred.wmt.arp.aic$pred + 2*pred.wmt.arp.aic$se)
legend(25,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast - AR(1)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt$pred - 2*pred.wmt$se)
lines(day_range[101:200], pred.wmt$pred + 2*pred.wmt$se)
legend(25,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast - AR(p)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp.arp.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp.arp.aic$pred - 2*pred.sp.arp.aic$se)
lines(day_range[101:200], pred.sp.arp.aic$pred + 2*pred.sp.arp.aic$se)
legend(25,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast - AR(1)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp$pred - 2*pred.sp$se)
lines(day_range[101:200], pred.sp$pred + 2*pred.sp$se)
legend(25,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
graphics.off()
```

# Fitting MA models using MA(3) and MA(q)

```{r}
fitPayxMa3 <- arima(df_log_ret[,"PAYX"], order = c(0,0,3))
fitPyplMa3 <- arima(df_log_ret[,"PYPL"], order = c(0,0,3))
fitWmtMa3 <- arima(df_log_ret[,"WMT"], order = c(0,0,3))
fitSpMa3 <- arima(df_log_ret[,"SP"], order = c(0,0,3))

fitPayxMaqAic <- auto.arima(df_log_ret[,"PAYX"], max.p = 0, max.q = 20, ic = "bic")
fitPyplMaqAic <- auto.arima(df_log_ret[,"PYPL"], max.p = 0, max.q = 20, ic = "bic")
fitWmtMaqAic <- auto.arima(df_log_ret[,"WMT"], max.p = 0, max.q = 20, ic = "bic")
fitSpMaqAic <- auto.arima(df_log_ret[,"SP"], max.p = 0, max.q = 20, ic = "bic")

Box.test(residuals(fitPayxMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPayxMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitPyplMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPyplMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitWmtMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitWmtMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitSpMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitSpMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)
```

```{r}
pred.payx.ma3 <- predict(fitPayxMa3, n.ahead = 100, se.fit = TRUE)
pred.pypl.ma3 <- predict(fitPyplMa3, n.ahead = 100, se.fit = TRUE)
pred.wmt.ma3 <- predict(fitWmtMa3, n.ahead = 100, se.fit = TRUE)
pred.sp.ma3 <- predict(fitSpMa3, n.ahead = 100, se.fit = TRUE)

pred.payx.maq.aic <- predict(fitPayxMaqAic, n.ahead = 100, se.fit = TRUE)
pred.pypl.maq.aic <- predict(fitPyplMaqAic, n.ahead = 100, se.fit = TRUE)
pred.wmt.maq.aic <- predict(fitWmtMaqAic, n.ahead = 100, se.fit = TRUE)
pred.sp.maq.aic <- predict(fitSpMaqAic, n.ahead = 100, se.fit = TRUE)
```

```{r}
pdf("Figure14-Forecast-Payx-Pypl_MA3_MAq.pdf",height=7,width=8)
par(mfrow=c(2,2))
#
plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast - MA(3)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx.ma3$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx.ma3$pred - 2*pred.payx.ma3$se)
lines(day_range[101:200], pred.payx.ma3$pred + 2*pred.payx.ma3$se)
legend(25,0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"PAYX"], type = "b", 
     main = "Paycheck Inc Forecast - MA(q)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.payx.maq.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.payx.maq.aic$pred - 2*pred.payx.maq.aic$se)
lines(day_range[101:200], pred.payx.maq.aic$pred + 2*pred.payx.maq.aic$se)
legend(25,0.001,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast - AR(p)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl.ma3$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl.ma3$pred - 2*pred.pypl.ma3$se)
lines(day_range[101:200], pred.pypl.ma3$pred + 2*pred.pypl.ma3$se)
legend(25,0.09,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"PYPL"], type = "b", 
     main = "Paypal Forecast - AR(1)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.pypl.maq.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.pypl.maq.aic$pred - 2*pred.pypl.maq.aic$se)
lines(day_range[101:200], pred.pypl.maq.aic$pred + 2*pred.pypl.maq.aic$se)
legend(25,0.09,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
graphics.off()
```

```{r}
pdf("Figure15-Forecast-Wmt-Sp_MA3_MAq.pdf",height=7,width=8)
par(mfrow=c(2,2))
#
plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast - MA(3)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt.ma3$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt.ma3$pred - 2*pred.wmt.ma3$se)
lines(day_range[101:200], pred.wmt.ma3$pred + 2*pred.wmt.ma3$se)
legend(25,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"WMT"], type = "b", 
     main = "Walmart Forecast - MA(q)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.wmt.maq.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.wmt.maq.aic$pred - 2*pred.wmt.maq.aic$se)
lines(day_range[101:200], pred.wmt.maq.aic$pred + 2*pred.wmt.maq.aic$se)
legend(25,0.06,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast - MA(3)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp.ma3$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp.ma3$pred - 2*pred.sp.ma3$se)
lines(day_range[101:200], pred.sp.ma3$pred + 2*pred.sp.ma3$se)
legend(25,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))

plot(day_range[1:100],xts1[t1,"SP"], type = "b", 
     main = "S&P 500 Forecast - MA(q)", ylab = "log_return", xlab = "Days", 
     xlim = c(-100,100))
points(day_range[101:200], pred.sp.maq.aic$pred,type="p",pch="*")
lines(day_range[101:200], pred.sp.maq.aic$pred - 2*pred.sp.maq.aic$se)
lines(day_range[101:200], pred.sp.maq.aic$pred + 2*pred.sp.maq.aic$se)
legend(25,0.1,c("data","predictions","lower CL","upper CL"),cex=0.9,
       box.lty=0,pch=c("o","*",NA,NA),lty=c(NA,NA,1,1))
#
graphics.off()
```


# Fitting ARMA models using ARMA(3,3) and ARMA(p,q)

```{r}
fitPayxArMa33 <- arima(df_log_ret[,"PAYX"], order = c(3,0,3))
fitPyplArMa33 <- arima(df_log_ret[,"PYPL"], order = c(3,0,3))
fitWmtArMa33 <- arima(df_log_ret[,"WMT"], order = c(3,0,3))
fitSpArMa33 <- arima(df_log_ret[,"SP"], order = c(3,0,3))

fitPayxArmapqAic <- auto.arima(df_log_ret[,"PAYX"], max.p = 20, max.q = 20, ic = "bic")
fitPyplArmapqAic <- auto.arima(df_log_ret[,"PYPL"], max.p = 20, max.q = 20, ic = "bic")
fitWmtArmapqAic <- auto.arima(df_log_ret[,"WMT"], max.p = 20, max.q = 20, ic = "bic")
fitSpArmapqAic <- auto.arima(df_log_ret[,"SP"], max.p = 20, max.q = 20, ic = "bic")

Box.test(residuals(fitPayxArma3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPayxMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitPyplMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitPyplMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitWmtMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitWmtMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)

Box.test(residuals(fitSpMa3), lag = 5, type = "Ljung-Box", fitdf = 1)
Box.test(residuals(fitSpMaqAic), lag = 5, type = "Ljung-Box", fitdf = 1)
```

